---
title: "Gtsummary in quarto"
format:
  html:
    theme: cosmo
    background-color: #11ffee00;
    fontcolor: "navy" 
    mainfont: sans-serif 
execute:
  echo: false
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(gtsummary)
#| echo: false
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd",
               "id", "nsibs", "samp", "race_eth", "sex", "region",
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"),
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |>
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes")))
```

# Customization of `tbl_summary()`
# Background:
 We’re using some data from the National Longitudinal Survey of Youth 1979, a cohort of American young adults aged 14-22 at enrollment in 1979. They continue to be followed to this day, and there is a wealth of publicly available data online. I’ve downloaded the answers to a survey question about whether respondents wear glasses, a scale about their eyesight with glasses, their (NLSY-assigned 😒) race/ethnicity, their sex (presumably at birth, but unclear), their family’s income in 1979, and their age at the birth of their first child.

# Table summarizing charactertistics of Young Adult

-race
-region 
-eyesight
-glasses
-age

```{r}
#| label: tbl-one
#| tbl-cap: "Table of characteritsics stratified by sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```
Figure 1: a table with the young adults characteritics stratified by sex

@tbl-one contains information about young adults characteritics 

# MORE TABLES 

```{r}
#| label: tbl-two
#| tbl-cap: "Table of characteritsics stratified by sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")

```
Figure 2: a table with the young adults characteritics stratified by sex



# MORE TABLES 

```{r}
#| label: tbl-three
#| tbl-cap: "Table of characteritsics stratified by sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```
Figure 3: a table with the young adults characteritics stratified by sex


# Quarto in class practice

```{r}
#| label: tbl-four
#| tbl-cap: "Table including amount of sleep stratified by sex"
tbl_four <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, sleep_wknd, sleep_wkdy, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir)) |>
  add_overall(last = TRUE)

tbl_four
```
Figure 4: a table with the young adults characteritics stratified by sex 


@tbl-four contains information and @tbl-one a table.

```{r}
stats <- list(
  n = nrow(nlsy),
  median_sleep_wknd = median(nlsy$sleep_wknd, na.rm = TRUE)
)

stats <- list(
  n = nrow(nlsy),
  median_sleep_wkdy = median(nlsy$sleep_wkdy, na.rm = TRUE)
)
```


```{r}
inline_text(tbl_four, variable = "sleep_wknd")


inline_text(tbl_four, variable = "sleep_wkdy")
```

This was from "Calculate the statistic directly as inline text from the table you created"

```{r}
inline_text(tbl_four, variable = "sleep_wknd", 
    pattern = "{median})")


inline_text(tbl_four, variable = "sleep_wkdy",
					    pattern = "{median})")
```
Then i was asked to Reformat that inline text from the table so that it only prints the median, not the IQR.


There were `r stats$n` participants with a median weekend sleep of `r stats$median_sleep_wknd` hours.

There were `r stats$n` participants with a median weekday sleep of `r stats$median_sleep_wkdy` hours.

**Add a chunk where you calculate median number of hours of sleep on weekends overall in the data (you can use the median(., na.rm = TRUE) function). Make sure your code prints that output when it renders.

How to Change the means to medians in your table and re-render?
```{r}

```



```



